TEMPERATURES AND METALLICITIES OF M DWARFS IN THE \apogee\ SURVEY

### 2019-07-31 paper/draft.tex
grants
abstract grammer

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
with their properties and compositions.
However, current atmospheric models
have limited abilities at reproducing spectral features in stars of the 
coolest temperature regimes ($T_{\rm eff} < 4200\,$K) and fully
exploiting the information content and statistical extent of
current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities,
spectral types for 5,875 M dwarfs in the \apogee\ survey using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K callibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex callibrated with FGK binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known SDSS optical spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
apply our models to a selection of 5,875 M dwarfs. 
We find \cannon\ temperatures to be in agreement to 60\,K
compared to a subsample of 1,702 sources with color-derived temperatures, and 
find \cannon\ metallicites to be in agreement to 0.08\,dex metallicity
compared to a subsample of 15 FGK+M or M+M binaries. 
Finally, our comparison between \cannon\ and APOGEE pipeline (ASPCAP) labels finds that 
ASPCAP is systematically biased towards reporting higher temperatures and lower metallicities. 
\end{abstract}

@@

%%==========================================================================================
\acknowledgements
We would like to acknowledge Hans Walter Rix (MPIA), Derek Homeier (Universit{\"a}t Heidelberg), Wolfgang Bradner (MPIA), Anna-Christina Eilers (MPIA), Melissa Ness (Columbia), Kevin Covey (WWU), Diogo Souto (Observatário Nacional/MCTI) and Keivan Stassun (Vanderbuilt) for constructive discussions in the process of this project, as well as Bertrand Goldman (MPIA) and those who have supported the internship program at the Max Planck Institute f{\"u}r Astronomie for providing JB with funding and hospitality.
This project was partially supported by
the US National Aeronautics and Space Administration (NASA grant NNX12AI50G),
the US National Science Foundation (NSF grant AST-1517237),
and the Moore--Sloan Data Science Environment at NYU.

@@

\color{gcolor}{ADAM: please add NASA and NSF grants in the parentheses above, and do you have a FAST acknowledgement?}\color{black}

>>>>

M dwarfs have enourmous potential for understanding structure and
formation on both Galactic and exoplanetary scales
through their properties and compositions.
However, current atmospheric models
have limited ability to reproduce spectral features in stars at the 
coolest temperatures ($T_{\rm eff} < 4200\,$K) and to fully
exploit the information content of current and upcoming large-scale spectroscopic surveys.
Here we present a catalog of spectroscopic temperatures, metallicities, and
spectral types for 5,875 M dwarfs in the \apogee+\drtwo\ surveys using
\thecannon: a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs with optically derived labels spanning 
$2860 < \teff < 4130$\,K calibrated with bolometric temperatures, 
$-0.5 < \feh < 0.5$\,dex calibrated with FGK binary metallicities, 
we train a two-parameter model with predictive accuracy (in cross-validation) 
to 77\,K and 0.09\,dex respectively.
We also train a one-dimensional spectral classification model using 51 M dwarfs
with SDSS optical spectral types ranging from M0 to M9, to predictive accuracy of
0.9 types. 
We find \cannon\ temperatures to be in agreement to within 60\,K
compared to a subsample of 1,702 sources with color-derived temperatures, and 
\cannon\ metallicites to be in agreement to within 0.08\,dex metallicity
compared to a subsample of 15 FGK+M or M+M binaries. 
Finally, our comparison between \cannon\ and APOGEE pipeline (ASPCAP) labels finds that 
ASPCAP is systematically biased towards reporting higher temperatures and lower metallicities. 
\end{abstract}

@@

%%==========================================================================================
\acknowledgements
We would like to acknowledge Hans Walter Rix (MPIA), Derek Homeier (Universit{\"a}t Heidelberg), Wolfgang Bradner (MPIA), Anna-Christina Eilers (MPIA), Melissa Ness (Columbia), Kevin Covey (WWU), Diogo Souto (Observatário Nacional/MCTI) and Keivan Stassun (Vanderbuilt) for constructive discussions in the process of this project, as well as Bertrand Goldman (MPIA) and those who have supported the internship program at the Max Planck Institute f{\"u}r Astronomie for providing JB with funding and hospitality.
This project was partially supported by
the US National Aeronautics and Space Administration (NASA grant NNX12AI50G),
the US National Science Foundation (NSF grant AST-1517237),
and the Moore--Sloan Data Science Environment at NYU.
A.J.B. acknowledges funding support from the National Science Foundation under award No. AST-1517177.
This work is supported by the SDSS Faculty and Student Team (FAST) initiative.

@@
